### 2023-03-15 main.tex
DualityNew

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}

\newcommand{\oad}{\mathbf{a}^{\dagger}}

@@

\section{Duality}
\subsection{Definition}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
\end{equation}
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
\newline
\newline
\textbf{Remark}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\end{equation}
\subsection{Duality for the multi-species stirring process}
We now state our duality result. The process $|\mathbf{n}\rangle$ with Hamiltonian \eqref{OriginalHamiltonian} is dual in the sense of \eqref{DualityRelation}, with respect to a process $|\bm{\xi}\rangle$ defined on the enlarged state space
\begin{equation}
    \widetilde{\Omega}:=\Omega\times \mathbb{N}_{0}^{V}
\end{equation}
with elements
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|r_{1}^{x},\ldots,r_{N}^{x}\rangle\otimes |m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}\rangle\right)
\end{equation}
where $r_{x}^{k}$ denotes the number of dual particles of type $k\in \{1,\ldots,N\}$ at site $x\in V$ and $m_{k}^{\widetilde{x}}$ denotes the number of particles of species $k\in \{2,\ldots,N\}$ at an extra-site $\widetilde{x}$ connected to the site $x\in V$.\\
The Hamiltonian of the dual process reads as
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}

where $H_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{k=2}^{N}\left((a^{\dagger})_{k}^{\widetilde{x}}\,E_{1k}^{x}-E_{kk}^{x}\right)
\end{equation}
By introducing the action of a bosonic creation operator $a^{\dagger}|q\rangle=|q+1\rangle$ on a generic vector $|q\rangle$ with $q\in \mathbb{N}_{0}$, we define in \eqref{siteDualHamiltonian} 
$(a^{\dagger})_{k}^{\widetilde{x}}$ as a copy of $a^{\dagger}$ acting on the extra-site $\widetilde{x}$ and on the occupation of the species $k\in\{2,\ldots,N\}$. \\
The Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual with respect to the duality matrix $D$ defined as 
\begin{equation}
\label{duality-function-good}
    D=\bigotimes_{x\in V}\left(D_{x}\otimes D_{\widetilde{x}}\right)
\end{equation}
We denote by $D_{x}$ the site part. This is defined as 
\begin{equation}
    D_{x}=R_{x}\exp{(E^{x})}
\end{equation}
with 
\begin{equation}
    E^{x}=\sum_{k=2}^{N}E_{k1}^{x}
\end{equation}
that implies 
\begin{equation}
    \exp{(E^{x})}=I+E^{x}
\end{equation}
and where $R_{x}$ is the diagonal matrix with elements (up to a finite constant) the inverse of the masses of the reversible measure \eqref{reversibleMeasure}, i.e. 
\begin{equation}
    R_{x}=\sum_{n\in\chi}\frac{\prod_{k=1}^{N}n_{k}^{x}}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\end{equation}
where, for all $x\in V$, we call $\chi=\{(n_{k}^{x})_{k\in\{1,\ldots,N\}}\in \mathbb{N}_{0}\,:\, \sum_{k=1}^{N}n_{k}^{x}=\nu\}$ the simplex.\\ 
We denote by $D_{\widetilde{x}}$ the extra-site part. It is defined as
\begin{equation}\label{dualityMatrix}
D_{\widetilde{x}}=\sum_{m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}=0}^{\infty}\prod_{k=2}^{N}\left(\frac{\alpha_{k}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)\langle m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}|
\end{equation}
On one hand, the edge part of the dual Hamiltonian \eqref{DualHamiltonian} is a copy of \eqref{edgeHamiltonian}, therefore it performs the stirring dynamic on the graph. On the other hand, the site dual Hamiltonian \eqref{siteDualHamiltonian} replaces a particle of any type $k\in\{2,\ldots,N\}$ at site $x$ with a particle of type $1$ and creates a particle of the same type $k$ at the extra-site $\widetilde{x}$. This means that, eventually, only particles of type $1$ remain in the graph while all other types are absorbed by the extra-sites. Thinking about the type $1$ as the hole  the graph is eventually emptied, similarly to what proved in the SEP(2j)\cite{carinci2013duality}.
\subsection{Proof of the duality relation}\label{subsectionDualityProof}
To prove the duality relation between \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} with the duality matrix \eqref{duality-function-good} we shall show that 
\begin{equation}
    H^{T}D=D\widetilde{H}
\end{equation}
Since the duality matrix is a tensor product over the sites $x\in V$ it is enough to prove that for the edge Hamiltonians
\begin{equation}\label{edgeDualRealtion}
    H_{x,y}^{T}D=DH_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and for the site Hamiltonians 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We will prove these two statements separately in the following.

\paragraph{Proof of \eqref{edgeDualRealtion}}
The proof of the edge duality relation relies on two facts: first the existence of a diagonal matrix, arising from the reversible measure, that is already a duality for the Hamiltonian (in literature called 'cheap' duality matrix, see \cite{giardina2009duality}); second the possibility of finding a proper symmetry of the Hamiltonian.  \\We first introduce the cheap duality matrix 

@@

R_{x}=\sum_{n\in\chi}\frac{\prod_{k=1}^{N}n_{k}^{x}}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
This is diagonal with the inverse of the masses of the reversible measure \eqref{reversibleMeasure}. 
This matrix connects the generators of the algebra $E^{x}_{ab}$ with their transposed $(E_{ab}^{x})^{T}$ by the following relation

@@

Indeed, we have
RE_{ab}^{x}R^{-1}=&\sum_{r\in\chi}\left(\frac{r_{1}^{x}!\ldots r_{N}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{1}^{x},\ldots, r_{N}^{x}|\right)

@@

	\sum_{s\in \chi,}\left(s_{b}^{x}|s_{1}^{x},\ldots,s_{a}^{x}+1,\ldots,s_{b}^{x}-1,\ldots s_{N}^{x}\rangle \langle s_{1}^{x},\ldots,s_{N}^{x}|\right)

@@

	\sum_{n\in\chi}\left(\frac{\nu!}{n_{1}!\ldots n_{N}!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle \langle n_{1}^{x},\ldots, n_{N}^{x}|\right)
	\sum_{r,n\in \chi}\frac{r_{1}^{x}!\ldots r_{N}^{x}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{0}^{x},\ldots, r_{N}^{x}|
	\frac{\nu!}{n_{1}^{x}!\ldots n_{N}^{x}!}n_{b}^{x}
 \qquad |n_{1}^{x},\ldots,n_{a}^{x}+1,\ldots,n_{b}^{x}-1,\ldots n_{N}^{x}\rangle \langle n_{1}^{x},\ldots,n_{N}^{x}|
 \sum_{r\in \chi}\frac{r_{1}^{x}!\ldots r_{N}^{x}!}{\nu!}\frac{\nu!}{r_{1}^{x}!\ldots (r_{a}^{x}-1)!\ldots(r_{b}^{x}+1)!\ldots r_{N}^{x}!}(r_{b}^{x}+1)
 \\&\qquad
 |r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{0}^{x},\ldots,r_{a}^{x}-1,\ldots,r_{b}^{x}-1,\ldots,r_{N}^{x}|
	\\=&\sum_{r\in\chi}\frac{r_{1}^{x}!\ldots r_{N}^{x}!}{\nu!}\frac{\nu!}{r_{1}^{x}!\ldots (r_{a}^{x}-1)!\ldots(r_{b}^{x}+1)!\ldots r_{N}^{x}!}(r_{b}^{x}+1)\frac{r_{a}^{x}}{r_{a}^{x}}
 \\& \qquad
 |r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{1}^{x},\ldots,r_{a}^{x}-1,\ldots,r_{b}^{x}-1,\ldots,r_{N}^{x}|
	\\=&\sum_{r\in \chi}

@@

where in the third to last equation we multiplied and divided for $r_{a}^{x}$. 
The relation \eqref{transpositionPropertyR} implies that 
    H_{x,y}^{T}=\left(R_{x}\otimes R_{y}\right)H_{x,y}\left(R_{x}\otimes R_{y}\right)^{-1}
We now look for a symmetry of the Hamiltonian. As shown in \eqref{hamiltonianCasimir}, it is possible to write the Hamiltonian \eqref{edgeHamiltonian} as a function of the coproduct of the second Casimir\eqref{secondCasimir} of $gl(N)$. We say that a matrix $A\in gl(N)$ is symmetry of $C_{2}$ if
    [A,C_{2}]=0
By the definition of the standard coproduct operator, it is possible to show that
    [A,C_{2}]=0\qquad \Longrightarrow \qquad \left[\Delta(A),\Delta(C_{2})\right]=0
 $C_{2}$ is a central element of $gl(N)$, then it commutes with each element of the algebra. 
We chose as a symmetry 

@@

    E^{x}=\sum_{k=2}E_{a1}^{x}
As a consequence of the previous considerations we can state that $E^{x}$ is a symmetry for the Hamiltonian \eqref{edgeHamiltonian}. We take the exponential of this symmetry by introducing
\begin{equation}\label{symmetry}
    S_{x,y}=\exp{(\Delta(E^{x}))}
This operator can be rewritten as 
		S&=e^{\Delta^{x,y}(E)}=\exp{(\mathbbm{1}^{x}\otimes E^{y}+E^{x}\otimes \mathbbm{1}^{y})}=\exp{(\mathbbm{1}^{x}\otimes E^{y})} \exp{(E^{x}\otimes \mathbbm{1}^{y})}\\&=\sum_{k=0}^{N}\frac{\left(\mathbbm{1}^{y}\otimes E^{x}\right)^{k}}{k!}\sum_{j=0}^{N}\frac{\left(E^{x}\otimes \mathbbm{1}^{y}\right)^{j}}{j!}=
		\left(\mathbbm{1}^{x}\otimes \sum_{k=0}^{N}\frac{(E^{y})^{k}}{k!}\right)\left(\sum_{j=0}^{N}\frac{(E^{x})^{j}}{j!}\otimes \mathbbm{1}^{y}\right)\\&=\left(\mathbbm{1}^{x}\otimes \exp{({E}^{y})}\right)\left(\exp{(E^{x})}\otimes \mathbbm{1}^{y}\right)\\&=\exp{(E^{x})}\otimes \exp{(E^{y})}
Moreover, by the exponential series of \eqref{equationE}, we have 
   \exp{(E^{x})}\otimes \exp{(E^{y})}= \left(\mathbbm{1}^{x}+E^{x}\right)\otimes \left(\mathbbm{1}^{y}+E^{y}\right)
Since \eqref{equationE} is a symmetry for $C_{2}$, then we have that
\begin{equation}\label{commutationSymm-edgeHamiltonian}
\left[H_{x,y},S\right]=\left[H_{x,y},\exp{(E^{x})}\otimes \exp{(E^{y})}\right]=0
By using \eqref{transpositionPropertyR} and \eqref{commutationSymm-edgeHamiltonian} we obtain
        H_{x,y}^{T}D=&\left(R_{x}\otimes R_{y}\right)H_{x,y}\left(R_{x}\otimes R_{y}\right)^{-1}D
        \left(R_{x}\otimes R_{y}\right)H_{x,y}\left(R_{x}\otimes R_{y}\right)^{-1}\left(R_{x}\otimes R_{y}\right)
        \left(\left(\exp{(E^{x})}\otimes D_{\widetilde{x}}\right)\otimes \left(\exp{(E^{y})}\otimes D_{\widetilde{y}}\right)\right)\bigotimes_{z\in V\,:\,z\neq x,y}(D_{z}\otimes D_{\widetilde{z}})
        \left(R_{x}\otimes R_{y}\right)H_{x,y}\left(\left(\exp{(E^{x})}\otimes D_{\widetilde{x}}\right)\otimes \left(\exp{(E^{y})}\otimes D_{\widetilde{y}}\right)\right)\bigotimes_{z\in V\,:\,z\neq x,y}(D_{z}\otimes D_{\widetilde{z}})

>>>>

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}
\newcommand{\dd}{\mathcal{D}_{\hat{x}}}
\newcommand{\oad}{\mathbf{a}^{\dagger}}

@@

\section{Duality}
\subsection{Definition}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{\,T}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\subsection{Duality for the multi-species stirring process}\label{statementDualitySubsection}
The process $(\mathbf{n}(t))_{t\geq 0}$, defined by the generator \eqref{Generator}, is dual to the absorbing boundary process $(\bm{\xi}(t))_{t\geq 0}$ living in the enlarged state space
\begin{equation}\label{dualStateSpace}
    \widetilde{\Omega}=\Omega_{x}^{V}\times \mathbb{N}_{0}^{V}
with $\forall x\in V$ and $\forall t\geq 0$ we define
    \xi^{x}(t)=\left(r_{1}^{x}(t),\ldots,r_{N}^{x}(t),m_{2}^{\hat{x}}(t),\ldots,m_{N}^{\hat{x}}(t)\right)
 where we denote $r_{k}^{x}(t)$ the occupation of dual particle of type $k\in \{1,\ldots,N\}$ at site $x$ at time $t\geq 0$ and where $m_{k}^{\hat{x}}(t)$ the occupation of dual particle of type $k\in \{1,\ldots,N\}$ at extra-site $\hat{x}$ connected to $x\in V$ at time $t\geq 0$.\\
 The dual generator is 
 \begin{equation}\label{DualGenerator}
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{k=2}^{N}\xi_{k}^{x}\left(f(\bm{\xi}-\delta_{k}^{x}+\delta_{1}^{x}+\delta_{k}^{\hat{x}})-f(\bm{\xi})\right)
and the duality function is 
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(2j-\sum_{k=2}^{N}r_{k}^{x})!}{\nu!}\prod_{k=2}^{N}\frac{n_{k}^{x}!}{(n_{k}^{x}-r_{k}^{x})!}\left(\frac{\alpha_{k}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)^{m_{k}^{\hat{x}}}\right)
On one hand, the edge part of the dual generator \eqref{DualGenerator} is a copy of \eqref{edgeGenerator}, therefore it performs the stirring dynamics on the graph. On the other hand, the site dual generator \eqref{siteDualGenerator} replaces a particle of any type $k\in\{2,\ldots,N\}$ at site $x$ with a particle of type $1$ and creates a particle of the same type $k$ at the extra-site $\hat{x}$. This last transition is performed with rate $\sum_{i=1}^{N}\alpha_{i}^{x}r_{k}^{x}$. This means that eventually the dual process empties the graph, putting all the dual particles of species $\{2,\ldots,N\}$ in the extra-sites, that play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ $\alpha_{k}^{x}=\alpha_{k}$, the average of the element of the duality matrix with respect to the reversible measure \eqref{reversibleMeasure}, is
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{k=2}^{N}\left(\frac{\alpha_{k}}{\sum_{i=1}^{N}\alpha_{i}}\right)^{\sum_{x\in V}r_{k}^{x}+\sum_{x\in V}m_{k}^{\hat{x}}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}

\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we shall show that \eqref{dualityRelationGenerator} is fulfilled. Here the state space is discrete, therefore the proof consists in finding an intertwining matrix $D$ that satisfies \eqref{dualityIntertwines}. To show this, we will use the Hamiltonians (linked with the generators by \eqref{Hamiltonian-Generato}) and their Lie algebraic description. Indeed, in this formalism the proof reduces in finding symmetries and exponential transformation of generators of the Lie algebra, that are the building blocks of the original and dual Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} respectively.  \\
The duality statement of subsection \ref{statementDualitySubsection} is rewritten in terms of Hamiltonian operators. The dual state space \eqref{dualStateSpace} is interpreted as a vector spaces with elements
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|r_{1}^{x},\ldots,r_{N}^{x}\rangle\otimes |m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}\rangle\right)
\end{equation}
The Hamiltonian of the dual process reads as
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}

where $H_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{k=2}^{N}\left((a^{\dagger})_{k}^{\hat{x}}\,E_{1k}^{x}-E_{kk}^{x}\right)
By introducing the action of a bosonic creation operator $a^{\dagger}|q\rangle=|q+1\rangle$ on a generic vector $|q\rangle$ with $q\in \mathbb{N}_{0}$, we define in \eqref{siteDualHamiltonian} 
$(a^{\dagger})_{k}^{\hat{x}}$ as a copy of $a^{\dagger}$ acting on the extra-site $\hat{x}$ and on the species $k\in\{2,\ldots,N\}$. \\
The Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual, in the sense of \eqref{DualityRelation} with respect to the duality matrix $D$ defined as 
\begin{equation}\label{dualityMatrix}
    D=\prod_{x\in V}d_{x}\otimes \dd
d_{x}=R_{x}\exp{(E^{x})}
\begin{equation}\label{EquationEx}
E^{x}=\sum_{a=2}^{N}E_{a1}^{x}
    R_{x}=\sum_{n^{x}\in\Omega_{x}}\frac{\prod_{k=1}^{N}n_{k}^{x}}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\begin{equation}\label{dualityMatrix}
\dd=\sum_{m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}=0}^{\infty}\prod_{k=2}^{N}\left(\frac{\alpha_{k}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)\langle m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}|
The matrix $R_{x}$ is diagonal. Their elements are related to the inverse of the weights of the reversible measure \eqref{reversibleMeasure}. In particular, to obtain these elements, we have considered the weights of \eqref{reversibleMeasure} when all the parameters $p_{i}=\frac{1}{N}$. Then, the constant $\left(\frac{1}{N}\right)^{\nu}$ has not been considered, since it does not change the duality relation. This $R_{x}$ is called the 'cheap' duality matrix (see \cite{giardina2009duality}). 
Since \eqref{dualityMatrix} is product over sites, \eqref{DualityRelation} is equivalent of proving that 
\begin{equation}\label{edgeDualRealtion}
    H_{x,y}^{T}D=DH_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
We perform the prove in three steps: first we will show that matrix \eqref{dualityMatrix} has indeed elements \eqref{dualityElements}; second we will prove \eqref{edgeDualRealtion}; finally we will show\eqref{siteDualRelation}. 
\paragraph{Proof of the elements of the duality matrix}We aim to show that 
\begin{equation}\label{proofDualityElements}
\langle \bm{n}|D|\bm{\xi}\rangle=D(\bm{n},\bm{\xi})\qquad   \forall \bm{n}\in \Omega,\quad \bm{\xi}\in \widetilde{\Omega}
with $D(\bm{n},\bm{\xi})$ defined in \eqref{dualityElements}. 
	 &\langle n^{x}|d_{x}\otimes \dd|\xi^{x}\rangle\\&=\langle n_{1}^{x},\ldots,n_{N}^{x}|\sum_{m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}=0}^{\infty}\prod_{k=2}^{N}\left(\frac{\alpha_{k}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)^{m_{k}^{\hat{x}}}\langle m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}|
	 \otimes(\exp{(E_{12}^{x}+\ldots+E_{1N}^{x}}))^{T}R_{x}\\&|r_{1}^{x},\ldots,r_{N}^{x}\rangle_{x} \otimes |q_{2}^{\hat{x}},\ldots,q_{N}^{\hat{x}}\rangle
On one hand, on the extra-site $\hat{x}$ we have 
\sum_{m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}=0}^{\infty}\prod_{k=2}^{N}\left(\frac{\alpha_{k}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)^{m_{k}^{\hat{x}}}\langle m_{2}^{\hat{x}},\ldots,m_{N}^{\hat{x}}||q_{2}^{\hat{x}},\ldots,q_{N}^{\hat{x}}|=\prod_{k=2}^{N}\left(\frac{\alpha_{k}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)^{m_{k}^{\hat{x}}}
where we used the orthogonality relation \eqref{ortho}. 
\frac{1}{k_{2}!,\ldots,k_{N}!}\frac{r_{1}^{x}!\ldots r_{N}^{x}!}{\nu!}|r_{1},\ldots,r_{N}\rangle_{x}\\&=
where in the last equality we applied the orthogonality relations \eqref{ortho} and the fact that $r_{1}^{x}=\nu-\sum_{k=2}^{N}r_{k}^{x}$. Finally, by taking the product over $x\in V$ \eqref{proofDualityElements} is proved.
\paragraph{Proof of \eqref{edgeDualRealtion}}To show \eqref{edgeDualRealtion} we need two ingredients. First is the existence of a similarity transformation between the Hamiltonian $H_{x,y}$ defined in \eqref{edgeHamiltonian} and its transposed. As we will show, this similarity transformation is $R_{x}$ defined in \eqref{Rmatrix}. Second, the possibility of finding a symmetry for the Hamiltonian \eqref{edgeHamiltonian}, exploiting the fact that $H_{x,y}$ is (up to a constant) the coproduct of the second Casimir of $gl(N)$, as shown in \eqref{hamiltonianCasimir}. As we will see, this symmetry is $exp{(E^{x})}\exp{(E^{y})}$ with $E^{x}$ defined in \eqref{equationEx}. \\
We show that 

@@

@@

Indeed
R_{x}E_{ab}^{x}R_{x}^{-1}=&\sum_{r^{x}\in\Omega_{x}}\left(\frac{r_{1}^{x}!\ldots r_{N}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{1}^{x},\ldots, r_{N}^{x}|\right)

@@

	\sum_{s^{x}\in \Omega_{x},}\left(s_{b}^{x}|s_{1}^{x},\ldots,s_{a}^{x}+1,\ldots,s_{b}^{x}-1,\ldots s_{N}^{x}\rangle \langle s_{1}^{x},\ldots,s_{N}^{x}|\right)

@@

	\sum_{n_{x}\in\Omega_{x}}\left(\frac{\nu!}{n_{1}!\ldots n_{N}!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle \langle n_{1}^{x},\ldots, n_{N}^{x}|\right)
 \\=&\sum_{r^{x}\in \Omega_{x}}

@@

where in the up to last equation we used the orthogonality relation \eqref{ortho}. 
Equation \eqref{transpositionPropertyR} implies that 
\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
We now search for a symmetry of $H_{x,y}$. Given $A,B\in gl(N)$, we say that $A$ is a symmetry of $B$ if 
    [A,B]=0
Moreover, since the coproduct is a Lie algebra homomorphism we have that 
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
Since $H_{x,y}$ is a linear function of the coproduct of the second Casimir, as a consequence of \eqref{symmetryCoproduct}, we can search for a symmetry of $C_{2}$. $C_{2}$ is a central for $gl(N)$, i.e. it commutes with all the element of the the algebra. We consider 

@@

    E=\sum_{a=2}^{N}E_{a2}
To obtain a product structure of the elements \eqref{dualityElements} of the duality matrix, we take the exponential of this matrix, i.e. we introduce the symmetry
    S=\exp{(\Delta(E))}
Since $[C_{2},E]=0$, we have that 
\begin{equation}\label{commutationSC}
    \left[S,\Delta(C_{2})\right]=0
For fixed $(x,y)\in \mathcal{E}$ and write
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. As a consequence of \eqref{commutationSC}, we have that 
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Finally, we have that 
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\hat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\hat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\hat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\hat{z}}\right)
        DH_{x,y}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 \newline\newline
 \textbf{Remark}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $k\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{k}$ is lower or equal than the number of original particles $n_{k}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing as a symmetry $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$. However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
